---
title: "Teaching"
site: distill::distill_website
---

## Workshop 1: Data Skills for Reproducible Science
[See](https://psyteachr.github.io/msc-data-skills/intro.html)


# Commands
Typing commands in the R Studio concole (bottom left window). Note that on a line in the editor, anything after is automagically green and treated as comment, not code (like '*' in Stata and SPSS)

```{r}
rm(list=ls())

1 + 1

1 + 2 + 3

1 +
  2
```



# Strings
```{r}
"Good morning"
```



Single quotes can be used as well, but not prefered
```{r}
'Good Morning'
```


# Quotes
Use single quotes if the string must cntain double quotes, or vice versa

```{r}
"'Good morning', she said"
'"Good morning", he replied'
```


# Cat
The command cat prints to the screen whatever is between brackets
```{r}
cat("'Good morning', she said")
cat('"Good morning", he replied')
```


# Variables
We start with variables that represent just one number (you may think of an SPSSS or Stata variable in a dataset with N=1)

Use the assignment operator '<-'

R stores the number in the variable 'variable1'
```{r}
variable1 <- 5
variable1
```

alternatively (seldom very useful)
```{r}
assign("variable2", 7)  # note the variable name is between quotes hre
variable2
```

We can use the variable in subsequent calculations 
```{r}
variable1 * variable2
# and assign the result to another variable
variable3 <- variable1 + variable2
```

Just as in SPSS, you can overwrite the original value
```{r}
variable1 <- variable1 * variable2
variable1
```



A variable does not need to be a single number, it can be a vector like our variables in Stata and SPSS

```{r}
myvariable <- c(2,3,5,7,11,13)
myvariable
```


Just as in Stata or SPSS, assignment of values to variables is static.
Here is an example from the online course:

```{r}
this_year <- 2019
my_birth_year <- 1976
my_age <- this_year - my_birth_year
this_year <- 2020
# your age is not updated automatically!
```

Look at the upper right panel ('Global Environment')

# Functions
Almost everything is R works with functions. hey are small programs that you feed with something and that return something else. 

Three numbers go in (the functionarguments), 10 come out (the value of the function)
```{r}
rnorm(n=10, mean=0, sd=1)
```

You often use a function in a function. sum() is also a function

```{r}
sum(rnorm(n=10, mean=0, sd=1))
```


Getting help is easy for beginners, but understanding it is not always easy

```{r}
help("rnorm")
?rnorm
```


# Packages
The power of R is in the packages. There are hundreds of them, for specific tasks. It can be overwhelming, esp. if multiple packages offer similar possibilities (i.e. inter-rater reliability assessment)

You can do it by commandline (take away #)

```{r}
# install.packages("ggExtra")
```

Or via menu (Tools-Install packages) or pane Packages (Install).
After installing: load package (one per session, typically at the top of yur script/syntax)

```{r}
library(ggExtra)
```

Read section 1.6.1 carefully. It is about how to organize your script
