---
title: "Conjunctive analysis of case configurations"
description: |
  What is conjunctive analysis and how to do it in R
author:
  - name: Asier Moneva
    url: https://nscr.nl/en/medewerker/asier-moneva/
date: 11-01-2022
output:
  distill::distill_article:
    self_contained: false

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this workshop **Asier Moneva** will explain what Conjunctive Analysis of Case Configurations (CACC) is about, and how it can be conducted in R with the help of the `cacc` package. Asier will first introduce the CACC and then perform a live demonstration of the `cacc` package from RStudio.  


Join this workshop meeting on Zoom by clicking [this link](https://us02web.zoom.us/j/81620041716?pwd=cjFCbG9TVHc0NTUrcmxmend5WHdCQT09)


> **Asier Moneva** is a postdoc at the NSCR and The Hague University of Applied Sciences with a background in criminology. He started specializing in crime analysis with R during his master's degree and has since developed an interest in open science.


### Materials

-   A [link](https://github.com/amoneva/cacc) to the development version of the `cacc` R package in GitHub. The `README.md` file contains a description of the package and walks readers through some examples.
-   [Placeholder for the slides from Asier's workshop presentation]

Asier will use the following R packages in the workshop:

-   `devtools`
-   `tidyverse`

To install them, you can run the code below:

```{r install}
#| message: false

if (! require("devtools")) install.packages(
  pkgs = "devtools",
  repos = "http://cran.us.r-project.org"
)

if (! require("tidyverse")) install.packages(
  pkgs = "tidyverse",
  repos = "http://cran.us.r-project.org"
)
```

